---
title: "NE25 Calibration Diagnostics"
subtitle: "Weight calibration quality for imputation `r params$imputation`"
author: "Kidsights Data Platform"
date: today
date-format: "YYYY-MM-DD HH:mm"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    fig-width: 9
    fig-height: 5
    fig-cap-location: bottom
    embed-resources: true
    smooth-scroll: true
execute:
  echo: false
  warning: false
  message: false
params:
  imputation: 1
---

```{r setup}
library(arrow)
library(dplyr)
library(ggplot2)
library(knitr)
library(tibble)

m <- params$imputation
# Paths are relative to this .qmd's location (Quarto default execute-dir = file)
data_root <- "../../../data/raking/ne25/ne25_weights"
diagnostics_file <- sprintf("%s/calibration_diagnostics_m%d.rds", data_root, m)
weights_file <- sprintf("%s/ne25_calibrated_weights_m%d.feather", data_root, m)

if (!file.exists(diagnostics_file)) {
  stop(sprintf("Diagnostics file not found: %s\nRun script 33 first.", diagnostics_file))
}
if (!file.exists(weights_file)) {
  stop(sprintf("Weights file not found: %s", weights_file))
}

diag <- readRDS(diagnostics_file)
weights_df <- arrow::read_feather(weights_file)

# Helper for formatting
fmt_pct <- function(x, digits = 2) sprintf("%.*f%%", digits, x)
fmt_num <- function(x, digits = 4) formatC(x, digits = digits, format = "f")
```

# Executive Summary

This report summarizes the quality of the calibrated raking weights produced by `scripts/raking/ne25/33_compute_kl_divergence_weights.R` for imputation **m = `r m`** of the NE25 analytic dataset. For the underlying method, see [`WEIGHT_CONSTRUCTION.qmd`](../../docs/raking/ne25/WEIGHT_CONSTRUCTION.qmd); for operator detail, see [`RAKING_INTEGRATION.md`](../../docs/raking/RAKING_INTEGRATION.md).

## Headline metrics

```{r headline, results='asis'}
# Support both new (post-fix) and legacy diagnostics files
stan_ok <- if (!is.null(diag$stan_terminated_normally)) diag$stan_terminated_normally else NA
within_1pct <- if (!is.null(diag$marginals_within_1pct)) diag$marginals_within_1pct else diag$converged

headline <- tibble::tibble(
  Metric = c(
    "Records calibrated (complete cases)",
    "Stan optimizer terminated normally",
    "All marginals within 1% of target",
    "Correlation RMSE — unweighted",
    "Correlation RMSE — weighted",
    "Correlation RMSE improvement",
    "Kish effective N",
    "Weight efficiency",
    "Weight ratio (max / min)"
  ),
  Value = c(
    formatC(diag$n_complete_cases, big.mark = ","),
    if (is.na(stan_ok)) "(not recorded — pre-fix diagnostics file)" else as.character(stan_ok),
    as.character(within_1pct),
    fmt_num(diag$unweighted_rmse_corr),
    fmt_num(diag$weighted_rmse_corr),
    fmt_pct(diag$correlation_improvement_pct, 1),
    formatC(round(diag$effective_n, 1), big.mark = ","),
    fmt_pct(diag$efficiency_pct, 1),
    fmt_num(diag$weight_ratio, 2)
  )
)
knitr::kable(headline, align = c("l", "r"))
```

::: {.callout-note}
**Two distinct convergence concepts.** `Stan optimizer terminated normally` indicates whether L-BFGS found a minimum cleanly (numerical convergence). `All marginals within 1%` indicates whether the solution satisfies the business-level calibration tolerance. These are orthogonal — Stan can converge to a minimum while some marginals remain outside 1% because of structural tension in the loss function (e.g., mental-health items at ~19% driven by the factorization, not a Stan failure). See [WEIGHT_CONSTRUCTION.qmd §3.3](../../docs/raking/ne25/WEIGHT_CONSTRUCTION.qmd) for the method and §5.2 for the mental-health discussion.
:::

# Weight Distribution

```{r weight-histogram, fig.cap = "Distribution of calibrated weights. Target = 1 (uniform); spread reflects how hard the optimizer had to push on individual records to match calibration targets."}
w <- weights_df$calibrated_weight

ggplot(tibble::tibble(w = w), aes(x = w)) +
  geom_histogram(bins = 60, fill = "#4c72b0", color = "white", alpha = 0.9) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", linewidth = 0.6) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 8)) +
  labs(
    x = "Calibrated weight",
    y = "Count",
    caption = sprintf(
      "N = %s · mean = %s · median = %s · max/min ratio = %s",
      formatC(length(w), big.mark = ","),
      fmt_num(mean(w), 3),
      fmt_num(median(w), 3),
      fmt_num(max(w) / min(w), 2)
    )
  ) +
  theme_minimal(base_size = 11)
```

```{r weight-quantiles, results='asis'}
q <- quantile(w, probs = c(0.01, 0.05, 0.25, 0.5, 0.75, 0.95, 0.99))
qtbl <- tibble::tibble(
  Quantile = c("1%", "5%", "25%", "50% (median)", "75%", "95%", "99%"),
  Weight = fmt_num(q, 3)
)
cat("## Weight quantiles\n\n")
knitr::kable(qtbl, align = c("l", "r"))
```

# Marginal Residuals

For each of the 24 calibration variables, the weighted sample mean should match the target mean. Residuals are shown in absolute terms and as a percentage of the target.

```{r marginal-residuals-table}
m_tbl <- diag$final_marginals %>%
  dplyr::mutate(
    Target = fmt_num(Target, 4),
    Achieved = fmt_num(Achieved, 4),
    Difference = fmt_num(Difference, 4),
    `Pct Diff` = fmt_pct(Pct_Diff, 2)
  ) %>%
  dplyr::select(Variable, Target, Achieved, Difference, `Pct Diff`, Pct_Diff) %>%
  dplyr::arrange(dplyr::desc(abs(Pct_Diff))) %>%
  dplyr::select(-Pct_Diff)

knitr::kable(m_tbl, align = c("l", "r", "r", "r", "r"))
```

```{r marginal-residual-plot, fig.cap = "Marginal residuals sorted by magnitude. Dashed lines mark 5% and 10% thresholds. Variables above 10% are structural flags warranting investigation."}
plt_df <- diag$final_marginals %>%
  dplyr::arrange(abs(Pct_Diff)) %>%
  dplyr::mutate(Variable = factor(Variable, levels = Variable))

ggplot(plt_df, aes(x = Variable, y = abs(Pct_Diff))) +
  geom_col(fill = "#55a868", alpha = 0.85) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#b07028", linewidth = 0.4) +
  geom_hline(yintercept = 10, linetype = "dashed", color = "red", linewidth = 0.4) +
  coord_flip() +
  labs(x = NULL, y = "|Pct Diff| (%)") +
  theme_minimal(base_size = 10)
```

# Correlation Matching

The primary calibration objective is matching the 488 observed cells of the 24 × 24 target covariance (correlation-scale loss). See [`WEIGHT_CONSTRUCTION.qmd §3.3`](../../docs/raking/ne25/WEIGHT_CONSTRUCTION.qmd#the-objective-function) for the formal loss.

## Summary

```{r corr-summary}
corr_tbl <- tibble::tibble(
  Metric = c(
    "Unweighted RMSE vs target correlation",
    "Weighted RMSE vs target correlation",
    "Absolute RMSE reduction",
    "Relative improvement"
  ),
  Value = c(
    fmt_num(diag$unweighted_rmse_corr),
    fmt_num(diag$weighted_rmse_corr),
    fmt_num(diag$unweighted_rmse_corr - diag$weighted_rmse_corr),
    fmt_pct(diag$correlation_improvement_pct, 2)
  )
)
knitr::kable(corr_tbl, align = c("l", "r"))
```

## Top correlation pairs by improvement

The 10 variable pairs where the weighted correlation moved closest to the target (largest error reduction).

```{r corr-top-improvements}
if (!is.null(diag$correlation_improvements_by_pair) && nrow(diag$correlation_improvements_by_pair) > 0) {
  imp_df <- diag$correlation_improvements_by_pair
  # Try common column names
  err_reduced_col <- intersect(c("error_reduced", "improvement", "error_reduction"), names(imp_df))
  if (length(err_reduced_col) > 0) {
    imp_df <- imp_df %>%
      dplyr::arrange(dplyr::desc(.data[[err_reduced_col[1]]])) %>%
      utils::head(10)
  }
  # Format numeric columns
  imp_df <- imp_df %>%
    dplyr::mutate(dplyr::across(dplyr::where(is.numeric), ~fmt_num(.x, 4)))
  knitr::kable(imp_df, align = c("l", "l", rep("r", ncol(imp_df) - 2)))
} else {
  cat("*(Correlation improvement detail not available in this diagnostics file.)*\n")
}
```

## Correlation matrix heatmaps

```{r corr-heatmaps, fig.width = 11, fig.height = 4.5, fig.cap = "Three 24×24 correlation matrices side by side. Top row labels correspond to calibration variables in the order they appear in the target moments structure."}
vars <- diag$variable_names

mats <- list(
  Unweighted = diag$unweighted_correlations,
  Target = diag$target_correlations,
  Weighted = diag$weighted_correlations
)

long_df <- purrr::map_dfr(names(mats), function(nm) {
  M <- mats[[nm]]
  dimnames(M) <- list(vars, vars)
  df <- as.data.frame(as.table(M))
  names(df) <- c("Var1", "Var2", "Correlation")
  df$Panel <- nm
  df
})

long_df$Panel <- factor(long_df$Panel, levels = c("Unweighted", "Target", "Weighted"))
long_df$Var1 <- factor(long_df$Var1, levels = vars)
long_df$Var2 <- factor(long_df$Var2, levels = rev(vars))

ggplot(long_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#3b5998", mid = "white", high = "#b23a48",
    midpoint = 0, limits = c(-1, 1)
  ) +
  facet_wrap(~Panel, nrow = 1) +
  coord_equal() +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6),
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

# Factorization & Block Structure

Reminder of the target covariance factorization (full detail in [`WEIGHT_CONSTRUCTION.qmd §3.3`](../../docs/raking/ne25/WEIGHT_CONSTRUCTION.qmd#factorized-target-structure)):

```{r block-table}
block_tbl <- tibble::tibble(
  Block = c("A — demographics", "B — PUMA", "C — mental health", "D — child outcome"),
  Vars = c(7, 14, 2, 1),
  `Identified from` = c(
    "Pooled ACS / NHIS / NSCH",
    "ACS only",
    "NHIS only",
    "NSCH only"
  ),
  `Effective n (per docs)` = c(
    "~2,000 (Block 1)",
    "~2,000 (Block 1)",
    "~1,500 (Block 2)",
    "~500 (Block 3)"
  )
)
knitr::kable(block_tbl, align = c("l", "r", "l", "l"))
```

The 88 structurally unidentified covariance cells (B × C = 28, B × D = 14, C × D = 2, doubled for both triangles) are masked out of the loss. See `cov_mask` in the unified moments for the exact pattern.

```{r pooling-weights, results='asis'}
if (!is.null(diag$pooling_weights)) {
  pw <- unlist(diag$pooling_weights)
  cat("## Pooling weights used for demographic targets\n\n")
  cat(sprintf(
    "- ACS: %s\n- NHIS: %s\n- NSCH: %s\n",
    fmt_pct(pw["acs"] * 100, 1),
    fmt_pct(pw["nhis"] * 100, 1),
    fmt_pct(pw["nsch"] * 100, 1)
  ))
}
```

# Reproducibility

- **Diagnostics RDS:** `r diagnostics_file`
- **Weights feather:** `r weights_file`
- **Rendered:** `r format(Sys.time(), "%Y-%m-%d %H:%M")`
- **R version:** `r R.version$version.string`

To regenerate:

```bash
# Refit weights for imputation m
"C:/Program Files/R/R-4.5.1/bin/Rscript.exe" \
  scripts/raking/ne25/33_compute_kl_divergence_weights.R

# Re-render this report for imputation m
"C:/Program Files/RStudio/resources/app/bin/quarto/bin/quarto.exe" \
  render scripts/raking/ne25/calibration_diagnostics_report.qmd \
  -P imputation:1
```
