---
title: "JSON Explorer"
subtitle: "Interactive navigation of the raw codebook data"
format:
  html:
    page-layout: full
    resources:
      - "assets/json-viewer.js"
      - "assets/prism.min.js"
      - "assets/prism.min.css"
---

```{r setup}
#| include: false
library(tidyverse)
library(jsonlite)
library(DT)
library(htmltools)

# Source codebook functions
source("../../R/codebook/load_codebook.R")
source("../../R/codebook/query_codebook.R")

# Load codebook
codebook <- load_codebook("../data/codebook.json", validate = FALSE)
```

<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/prism/1.29.0/themes/prism.min.css">
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/jstree/3.3.12/themes/default/style.min.css">

<style>
.json-explorer-container {
  display: flex;
  height: 70vh;
  border: 1px solid #ddd;
  border-radius: 8px;
  overflow: hidden;
  margin: 20px 0;
}

.tree-panel {
  width: 40%;
  background: #f8f9fa;
  border-right: 1px solid #ddd;
  overflow-y: auto;
  padding: 15px;
}

.json-panel {
  width: 60%;
  background: white;
  overflow-y: auto;
  padding: 15px;
}

.search-container {
  margin-bottom: 20px;
}

.search-input {
  width: 100%;
  padding: 10px;
  border: 1px solid #ced4da;
  border-radius: 4px;
  font-size: 16px;
}

.breadcrumb-nav {
  background: #e9ecef;
  padding: 8px 12px;
  border-radius: 4px;
  margin-bottom: 15px;
  font-family: monospace;
  font-size: 12px;
}

.json-viewer {
  font-family: 'Courier New', monospace;
  font-size: 13px;
  line-height: 1.4;
}

.json-key {
  color: #d63384;
  font-weight: bold;
}

.json-string {
  color: #198754;
}

.json-number {
  color: #0d6efd;
}

.json-boolean {
  color: #fd7e14;
}

.json-null {
  color: #6c757d;
  font-style: italic;
}

.copy-btn {
  position: absolute;
  top: 10px;
  right: 10px;
  background: #007bff;
  color: white;
  border: none;
  padding: 5px 10px;
  border-radius: 3px;
  cursor: pointer;
  font-size: 12px;
}

.copy-btn:hover {
  background: #0056b3;
}

.item-summary {
  background: #f1f3f4;
  padding: 15px;
  border-radius: 6px;
  margin-bottom: 15px;
  border-left: 4px solid #007bff;
}

.expand-toggle {
  cursor: pointer;
  user-select: none;
  color: #007bff;
  margin-right: 5px;
}

.expand-toggle:hover {
  color: #0056b3;
}

.collapsed-content {
  display: none;
}

.filters-panel {
  background: #fff;
  border: 1px solid #dee2e6;
  border-radius: 6px;
  padding: 15px;
  margin-bottom: 20px;
}

.filter-group {
  display: inline-block;
  margin-right: 15px;
  margin-bottom: 10px;
}

.filter-group select {
  padding: 5px 10px;
  border: 1px solid #ced4da;
  border-radius: 4px;
}

.item-grid {
  display: grid;
  grid-template-columns: repeat(auto-fill, minmax(400px, 1fr));
  gap: 15px;
  margin-top: 20px;
}

.item-card {
  border: 1px solid #dee2e6;
  border-radius: 8px;
  background: white;
  overflow: hidden;
  transition: box-shadow 0.2s ease;
}

.item-card:hover {
  box-shadow: 0 4px 12px rgba(0,0,0,0.1);
}

.item-header {
  background: linear-gradient(135deg, #007bff, #0056b3);
  color: white;
  padding: 12px 15px;
  cursor: pointer;
}

.item-body {
  padding: 15px;
}

.study-tags {
  margin-top: 8px;
}

.study-tag {
  display: inline-block;
  background: #e9ecef;
  color: #495057;
  padding: 2px 8px;
  border-radius: 12px;
  font-size: 11px;
  margin-right: 5px;
  margin-top: 3px;
}

.json-content {
  position: relative;
  background: #f8f9fa;
  border-radius: 4px;
  margin-top: 10px;
}

#json-tree {
  max-height: 50vh;
  overflow-y: auto;
}

.jstree-default .jstree-clicked {
  background: #007bff !important;
  color: white !important;
}
</style>

## Interactive JSON Explorer

Navigate through the raw codebook JSON data using multiple interfaces:

### Quick Search

```{r search-input}
#| results: asis
#| echo: false

cat('<div class="search-container">')
cat('<input type="text" id="global-search" class="search-input" placeholder="Search items by ID, text content, or domain..." autocomplete="off">')
cat('</div>')
```

### Filters

```{r filters}
#| results: asis
#| echo: false

# Create domains list using dataframe approach
df <- items_to_dataframe(codebook)
kidsights_domains <- unique(df$domain_kidsights)
cahmi_domains <- unique(df$domain_cahmi)
domains <- unique(c(kidsights_domains, cahmi_domains))
domains <- domains[!is.na(domains)]

# Create studies list
studies <- unique(unlist(sapply(codebook$items, function(x) x$studies)))

cat('<div class="filters-panel">')

# Domain filter
cat('<div class="filter-group">')
cat('<label for="domain-filter">Domain:</label>')
cat('<select id="domain-filter">')
cat('<option value="">All Domains</option>')
for(domain in sort(domains)) {
  cat('<option value="', domain, '">', domain, '</option>')
}
cat('</select>')
cat('</div>')

# Study filter
cat('<div class="filter-group">')
cat('<label for="study-filter">Study:</label>')
cat('<select id="study-filter">')
cat('<option value="">All Studies</option>')
for(study in sort(studies)) {
  cat('<option value="', study, '">', study, '</option>')
}
cat('</select>')
cat('</div>')

# View mode filter
cat('<div class="filter-group">')
cat('<label for="view-mode">View Mode:</label>')
cat('<select id="view-mode">')
cat('<option value="tree">Tree Navigator</option>')
cat('<option value="cards">Item Cards</option>')
cat('<option value="table">Data Table</option>')
cat('</select>')
cat('</div>')

cat('<button id="clear-filters" class="btn btn-sm btn-outline-secondary">Clear Filters</button>')
cat('</div>')
```

### Tree Navigator

```{r tree-navigator}
#| results: asis
#| echo: false

cat('<div id="tree-view" class="json-explorer-container">')
cat('<div class="tree-panel">')
cat('<h6>Codebook Structure</h6>')
cat('<div id="json-tree"></div>')
cat('</div>')

cat('<div class="json-panel">')
cat('<div class="breadcrumb-nav" id="breadcrumb">')
cat('Click an item in the tree to view its JSON structure')
cat('</div>')
cat('<div id="json-display" class="json-viewer">')
cat('<div class="item-summary">')
cat('<h5>Welcome to JSON Explorer</h5>')
cat('<p>Use the tree on the left to navigate through the codebook structure. Click on any item to see its complete JSON representation with syntax highlighting.</p>')
cat('<p><strong>Features:</strong></p>')
cat('<ul>')
cat('<li>🔍 Search through all items</li>')
cat('<li>🌳 Tree navigation of JSON structure</li>')
cat('<li>📋 Copy JSON to clipboard</li>')
cat('<li>🎨 Syntax highlighting</li>')
cat('<li>📱 Responsive design</li>')
cat('</ul>')
cat('</div>')
cat('</div>')
cat('</div>')
cat('</div>')
```

### Item Cards View

```{r cards-view}
#| results: asis
#| echo: false

cat('<div id="cards-view" style="display: none;">')
cat('<div id="item-cards" class="item-grid">')
cat('<!-- Cards will be populated by JavaScript -->')
cat('</div>')
cat('</div>')
```

### Data Table View

```{r table-view-container}
#| results: asis
#| echo: false

cat('<div id="table-view" style="display: none;">')
```

```{r interactive-table}
# Create comprehensive data frame for table view
df <- items_to_dataframe(codebook, flatten_identifiers = TRUE)

# Add JSON preview column
df$json_preview <- sapply(names(codebook$items), function(item_id) {
  paste0('<button class="btn btn-sm btn-outline-primary view-json" data-item="',
         item_id, '">View JSON</button>')
})

datatable(
  df %>%
    select(
      `Item ID` = item_id,
      `Domain (Kidsights)` = domain_kidsights,
      `Domain (CAHMI)` = domain_cahmi,
      `Studies` = studies,
      `Text` = stem_combined,
      `JSON` = json_preview
    ),
  filter = 'top',
  escape = FALSE,
  options = list(
    pageLength = 15,
    scrollX = TRUE,
    searchHighlight = TRUE,
    dom = 'Bfrtip',
    buttons = list(
      'copy',
      list(extend = 'csv', filename = paste0('codebook_table_', Sys.Date())),
      'colvis'
    ),
    columnDefs = list(
      list(width = '100px', targets = 0),  # Item ID
      list(width = '80px', targets = 1),   # Domain (Kidsights)
      list(width = '80px', targets = 2),   # Domain (CAHMI)
      list(width = '120px', targets = 3),  # Studies
      list(width = '300px', targets = 4),  # Text
      list(width = '80px', targets = 5)    # JSON button
    )
  ),
  extensions = 'Buttons'
) %>%
  formatStyle(
    'Domain (Kidsights)',
    backgroundColor = styleEqual(
      c('motor', 'socemo', 'coglan'),
      c('#4ECDC4', '#FF6B6B', '#45B7D1')
    ),
    color = 'white',
    fontWeight = 'bold'
  ) %>%
  formatStyle(
    'Domain (CAHMI)',
    backgroundColor = styleEqual(
      c('motor', 'early_learning', 'self_regulation', 'social_emotional'),
      c('#4ECDC4', '#96CEB4', '#FECA57', '#FF9999')
    ),
    color = 'white',
    fontWeight = 'bold'
  )
```

```{r close-table-view}
#| results: asis
#| echo: false

cat('</div>')
```

```{r modal-html}
#| results: asis
#| echo: false

# Modal for JSON display
cat('<!-- Modal for JSON display -->')
cat('<div class="modal fade" id="jsonModal" tabindex="-1">')
cat('<div class="modal-dialog modal-xl">')
cat('<div class="modal-content">')
cat('<div class="modal-header">')
cat('<h5 class="modal-title" id="modalTitle">Item JSON</h5>')
cat('<button type="button" class="btn-close" data-bs-dismiss="modal"></button>')
cat('</div>')
cat('<div class="modal-body">')
cat('<div style="position: relative;">')
cat('<button id="copy-modal-json" class="copy-btn">Copy JSON</button>')
cat('<pre id="modal-json-content" class="json-viewer"></pre>')
cat('</div>')
cat('</div>')
cat('</div>')
cat('</div>')
cat('</div>')
```

<script src="https://cdnjs.cloudflare.com/ajax/libs/jquery/3.6.0/jquery.min.js"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/jstree/3.3.12/jstree.min.js"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/prism/1.29.0/components/prism-core.min.js"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/prism/1.29.0/components/prism-json.min.js"></script>

```{r embed-codebook}
#| results: asis
# Embed the codebook JSON as a JavaScript variable
cat("<script type='application/json' id='codebook-data'>")
cat(toJSON(codebook, auto_unbox = TRUE, pretty = TRUE))
cat("</script>")
```

<script>
// Load codebook data
const codebookData = JSON.parse(document.getElementById('codebook-data').textContent);

// Initialize the application
$(document).ready(function() {
  initializeExplorer();
  populateFilters();
  setupEventHandlers();
});

function initializeExplorer() {
  // Initialize tree view
  initializeTree();

  // Show tree view by default
  $('#view-mode').val('tree');
  showView('tree');
}

function initializeTree() {
  const treeData = buildTreeData(codebookData);

  $('#json-tree').jstree({
    'core': {
      'data': treeData,
      'themes': {
        'responsive': true,
        'dots': true
      }
    },
    'plugins': ['search', 'wholerow']
  });

  // Handle tree selection
  $('#json-tree').on('select_node.jstree', function(e, data) {
    const node = data.node;
    const path = getNodePath(node);
    const value = getValueFromPath(codebookData, path);

    displayJSON(value, path);
  });
}

function buildTreeData(data) {
  const result = [];

  // Add metadata
  result.push({
    id: 'metadata',
    text: `Metadata (v${data.metadata.version})`,
    children: buildObjectChildren(data.metadata, 'metadata'),
    state: { opened: false }
  });

  // Add items (flat structure matching JSON)
  const itemsNode = {
    id: 'items',
    text: `Items (${Object.keys(data.items).length})`,
    children: [],
    state: { opened: true }
  };

  // Add items in their natural JSON order (no domain grouping)
  Object.keys(data.items).forEach(itemId => {
    const item = data.items[itemId];
    itemsNode.children.push({
      id: `items.${itemId}`,
      text: `${itemId} - ${item.content?.stems?.combined?.substring(0, 50) || 'No text'}...`,
      children: buildObjectChildren(item, `items.${itemId}`)
    });
  });

  result.push(itemsNode);

  // Add response sets
  if (data.response_sets) {
    result.push({
      id: 'response_sets',
      text: `Response Sets (${Object.keys(data.response_sets).length})`,
      children: buildObjectChildren(data.response_sets, 'response_sets'),
      state: { opened: false }
    });
  }

  // Add domains
  if (data.domains) {
    result.push({
      id: 'domains',
      text: `Domains (${Object.keys(data.domains).length})`,
      children: buildObjectChildren(data.domains, 'domains'),
      state: { opened: false }
    });
  }

  return result;
}

function buildObjectChildren(obj, path) {
  const children = [];

  if (Array.isArray(obj)) {
    obj.forEach((item, index) => {
      const value = typeof item === 'object' ? `{...}` : item;
      children.push({
        id: `${path}.${index}`,
        text: `[${index}]: ${value}`,
        children: typeof item === 'object' ? buildObjectChildren(item, `${path}.${index}`) : false
      });
    });
  } else if (typeof obj === 'object' && obj !== null) {
    Object.keys(obj).forEach(key => {
      const value = obj[key];
      const displayValue = typeof value === 'object' && value !== null
        ? (Array.isArray(value) ? `[...] (${value.length})` : `{...}`)
        : value;

      children.push({
        id: `${path}.${key}`,
        text: `${key}: ${displayValue}`,
        children: typeof value === 'object' && value !== null ? buildObjectChildren(value, `${path}.${key}`) : false
      });
    });
  }

  return children;
}

function getNodePath(node) {
  return node.id.split('.');
}

function getValueFromPath(data, path) {
  let current = data;

  for (let i = 0; i < path.length; i++) {
    const key = path[i];
    if (current && typeof current === 'object' && key in current) {
      current = current[key];
    } else {
      return null;
    }
  }

  return current;
}

function displayJSON(value, path) {
  const breadcrumb = path.join(' → ');
  $('#breadcrumb').text(breadcrumb);

  const jsonString = JSON.stringify(value, null, 2);
  const highlighted = Prism.highlight(jsonString, Prism.languages.json, 'json');

  $('#json-display').html(`
    <div style="position: relative;">
      <button class="copy-btn" onclick="copyToClipboard('${jsonString.replace(/'/g, "\\'")}')">Copy JSON</button>
      <pre class="language-json"><code>${highlighted}</code></pre>
    </div>
  `);
}

function populateFilters() {
  // Populate domain filter
  const domains = new Set();
  Object.values(codebookData.items).forEach(item => {
    if (item.classification?.domain) {
      domains.add(item.classification.domain);
    }
  });

  const domainSelect = $('#domain-filter');
  Array.from(domains).sort().forEach(domain => {
    domainSelect.append(`<option value="${domain}">${domain}</option>`);
  });

  // Populate study filter
  const studies = new Set();
  Object.values(codebookData.items).forEach(item => {
    if (item.studies && Array.isArray(item.studies)) {
      item.studies.forEach(study => studies.add(study));
    }
  });

  const studySelect = $('#study-filter');
  Array.from(studies).sort().forEach(study => {
    studySelect.append(`<option value="${study}">${study}</option>`);
  });
}

function setupEventHandlers() {
  // Search functionality
  $('#global-search').on('keyup', function() {
    const searchTerm = $(this).val();

    if (searchTerm.length > 2) {
      $('#json-tree').jstree('search', searchTerm);
      filterItems(searchTerm);
    } else {
      $('#json-tree').jstree('clear_search');
      showAllItems();
    }
  });

  // View mode switching
  $('#view-mode').on('change', function() {
    const mode = $(this).val();
    showView(mode);
  });

  // Filter changes
  $('#domain-filter, #study-filter').on('change', function() {
    applyFilters();
  });

  // Clear filters
  $('#clear-filters').on('click', function() {
    $('#domain-filter, #study-filter').val('');
    $('#global-search').val('');
    $('#json-tree').jstree('clear_search');
    showAllItems();
  });

  // Modal JSON view buttons
  $(document).on('click', '.view-json', function() {
    const itemId = $(this).data('item');
    const item = codebookData.items[itemId];

    $('#modalTitle').text(`Item: ${itemId}`);
    const jsonString = JSON.stringify(item, null, 2);
    const highlighted = Prism.highlight(jsonString, Prism.languages.json, 'json');
    $('#modal-json-content').html(highlighted);

    $('#jsonModal').modal('show');
  });

  // Copy modal JSON
  $('#copy-modal-json').on('click', function() {
    const jsonText = $('#modal-json-content').text();
    copyToClipboard(jsonText);
  });
}

function showView(mode) {
  $('#tree-view, #cards-view, #table-view').hide();

  switch(mode) {
    case 'tree':
      $('#tree-view').show();
      break;
    case 'cards':
      $('#cards-view').show();
      generateItemCards();
      break;
    case 'table':
      $('#table-view').show();
      break;
  }
}

function generateItemCards() {
  const container = $('#item-cards');
  container.empty();

  Object.keys(codebookData.items).forEach(itemId => {
    const item = codebookData.items[itemId];
    const card = createItemCard(itemId, item);
    container.append(card);
  });
}

function createItemCard(itemId, item) {
  const studies = item.studies ? item.studies.map(s => `<span class="study-tag">${s}</span>`).join('') : '';
  const domain = item.classification?.domain || 'unknown';
  const text = item.content?.stems?.combined || 'No text available';

  return $(`
    <div class="item-card" data-item="${itemId}">
      <div class="item-header" onclick="toggleCard('${itemId}')">
        <h6 class="mb-1">${itemId}</h6>
        <small>${domain}</small>
        <div class="study-tags">${studies}</div>
      </div>
      <div class="item-body">
        <p class="text-muted small">${text.substring(0, 200)}...</p>
        <div class="json-content collapsed-content" id="json-${itemId}">
          <button class="copy-btn" onclick="copyItemJSON('${itemId}')">Copy</button>
          <pre class="language-json"><code>${Prism.highlight(JSON.stringify(item, null, 2), Prism.languages.json, 'json')}</code></pre>
        </div>
      </div>
    </div>
  `);
}

function toggleCard(itemId) {
  $(`#json-${itemId}`).toggleClass('collapsed-content');
}

function copyItemJSON(itemId) {
  const item = codebookData.items[itemId];
  const jsonString = JSON.stringify(item, null, 2);
  copyToClipboard(jsonString);
}

function copyToClipboard(text) {
  navigator.clipboard.writeText(text).then(function() {
    // Could add a toast notification here
    console.log('JSON copied to clipboard');
  });
}

function filterItems(searchTerm) {
  // Implementation for filtering items based on search
  // This would filter the tree, cards, and table views
}

function applyFilters() {
  const domain = $('#domain-filter').val();
  const study = $('#study-filter').val();

  // Implementation for applying domain and study filters
}

function showAllItems() {
  // Implementation for showing all items (removing filters)
}
</script>