---
title: "Visual Overview"
subtitle: "Interactive visualizations of codebook structure and relationships"
format:
  html:
    page-layout: full
---

```{r setup}
#| include: false
library(tidyverse)
library(jsonlite)
library(plotly)
library(networkD3)
library(visNetwork)

# Source codebook functions
source("../../R/codebook/load_codebook.R")
source("../../R/codebook/query_codebook.R")

# Load codebook
codebook <- load_codebook("../data/codebook.json", validate = FALSE)
```

<style>
.viz-container {
  border: 1px solid #dee2e6;
  border-radius: 8px;
  padding: 20px;
  margin: 20px 0;
  background: white;
  box-shadow: 0 2px 4px rgba(0,0,0,0.1);
}

.viz-controls {
  background: #f8f9fa;
  padding: 15px;
  border-radius: 6px;
  margin-bottom: 15px;
}

.control-group {
  display: inline-block;
  margin-right: 20px;
  margin-bottom: 10px;
}

.control-group label {
  display: block;
  font-weight: 500;
  margin-bottom: 5px;
  font-size: 14px;
}

.control-group select, .control-group input {
  padding: 6px 10px;
  border: 1px solid #ced4da;
  border-radius: 4px;
  font-size: 14px;
}

.network-info {
  background: #e7f3ff;
  border: 1px solid #b3d9ff;
  border-radius: 4px;
  padding: 10px;
  margin-bottom: 15px;
  font-size: 14px;
}

.legend {
  display: flex;
  flex-wrap: wrap;
  gap: 15px;
  margin-top: 10px;
  padding: 10px;
  background: #f8f9fa;
  border-radius: 4px;
}

.legend-item {
  display: flex;
  align-items: center;
  gap: 5px;
  font-size: 13px;
}

.legend-color {
  width: 16px;
  height: 16px;
  border-radius: 3px;
}

.stats-grid {
  display: grid;
  grid-template-columns: repeat(auto-fit, minmax(200px, 1fr));
  gap: 15px;
  margin: 20px 0;
}

.stat-card {
  background: linear-gradient(135deg, #667eea 0%, #764ba2 100%);
  color: white;
  padding: 20px;
  border-radius: 8px;
  text-align: center;
}

.stat-number {
  font-size: 2.5em;
  font-weight: bold;
  line-height: 1;
}

.stat-label {
  font-size: 0.9em;
  opacity: 0.9;
  margin-top: 5px;
}

#sunburst-chart {
  height: 600px;
}

#network-chart {
  height: 700px;
  border: 1px solid #ddd;
}

#treemap-chart {
  height: 500px;
}

.tooltip {
  position: absolute;
  background: rgba(0,0,0,0.8);
  color: white;
  padding: 8px 12px;
  border-radius: 4px;
  font-size: 12px;
  pointer-events: none;
  z-index: 1000;
}
</style>

## Overview Statistics

```{r overview-stats}
#| results: asis
# Calculate key statistics
total_items <- length(codebook$items)
total_studies <- length(unique(unlist(sapply(codebook$items, function(x) x$studies))))
total_domains <- length(unique(sapply(codebook$items, function(x) x$classification$domain)))

# Items with response options
items_with_responses <- sum(sapply(codebook$items, function(x) {
  !is.null(x$content$response_options) && length(x$content$response_options) > 0
}))

# Calibration items
calibration_items <- sum(sapply(codebook$items, function(x) {
  isTRUE(x$psychometric$calibration_item)
}))

cat('<div class="stats-grid">')
cat('<div class="stat-card"><div class="stat-number">', total_items, '</div><div class="stat-label">Total Items</div></div>')
cat('<div class="stat-card"><div class="stat-number">', total_studies, '</div><div class="stat-label">Studies</div></div>')
cat('<div class="stat-card"><div class="stat-number">', total_domains, '</div><div class="stat-label">Domains</div></div>')
cat('<div class="stat-card"><div class="stat-number">', items_with_responses, '</div><div class="stat-label">Items with Responses</div></div>')
cat('<div class="stat-card"><div class="stat-number">', calibration_items, '</div><div class="stat-label">Calibration Items</div></div>')
cat('</div>')
```

## Interactive Hierarchical View

### Sunburst Chart - Navigate the Codebook Hierarchy

<div class="viz-container">
  <div class="viz-controls">
    <div class="control-group">
      <label for="sunburst-level">Detail Level:</label>
      <select id="sunburst-level">
        <option value="1">Domain Level</option>
        <option value="2">Domain → Items</option>
        <option value="3">Domain → Study → Items</option>
      </select>
    </div>
    <div class="control-group">
      <label for="sunburst-metric">Size Metric:</label>
      <select id="sunburst-metric">
        <option value="count">Item Count</option>
        <option value="studies">Study Coverage</option>
      </select>
    </div>
  </div>

  <div id="sunburst-chart"></div>

  <div class="legend">
    <div class="legend-item">
      <div class="legend-color" style="background: #FF6B6B;"></div>
      <span>Social-Emotional</span>
    </div>
    <div class="legend-item">
      <div class="legend-color" style="background: #4ECDC4;"></div>
      <span>Motor</span>
    </div>
    <div class="legend-item">
      <div class="legend-color" style="background: #45B7D1;"></div>
      <span>Cognitive-Language</span>
    </div>
    <div class="legend-item">
      <div class="legend-color" style="background: #96CEB4;"></div>
      <span>Early Learning</span>
    </div>
    <div class="legend-item">
      <div class="legend-color" style="background: #FECA57;"></div>
      <span>Self-Regulation</span>
    </div>
  </div>
</div>

## Network Visualization

### Item-Study Relationships

<div class="viz-container">
  <div class="network-info">
    <strong>📊 Network View:</strong> This interactive network shows how items (circles) connect to studies (squares).
    Hover over nodes to see details, click and drag to explore. Use the controls below to filter and customize the view.
  </div>

  <div class="viz-controls">
    <div class="control-group">
      <label for="network-filter-domain">Filter by Domain:</label>
      <select id="network-filter-domain">
        <option value="">All Domains</option>
      </select>
    </div>

    <div class="control-group">
      <label for="network-layout">Layout:</label>
      <select id="network-layout">
        <option value="forceAtlas2Based">Force Atlas</option>
        <option value="barnesHut">Barnes Hut</option>
        <option value="repulsion">Repulsion</option>
        <option value="hierarchicalRepulsion">Hierarchical</option>
      </select>
    </div>

    <div class="control-group">
      <label for="show-labels">Show Labels:</label>
      <input type="checkbox" id="show-labels" checked>
    </div>

    <div class="control-group">
      <label for="node-size">Node Size by:</label>
      <select id="node-size">
        <option value="connections">Connections</option>
        <option value="uniform">Uniform</option>
      </select>
    </div>
  </div>

  <div id="network-chart"></div>
</div>

## Domain Analysis

### Treemap - Items by Domain and Study

<div class="viz-container">
  <div class="viz-controls">
    <div class="control-group">
      <label for="treemap-grouping">Grouping:</label>
      <select id="treemap-grouping">
        <option value="domain-study">Domain → Study</option>
        <option value="study-domain">Study → Domain</option>
        <option value="domain-only">Domain Only</option>
      </select>
    </div>
  </div>

  <div id="treemap-chart"></div>
</div>

```{r embed-data-for-js}
#| results: asis
# Prepare data for JavaScript visualizations
df <- items_to_dataframe(codebook, flatten_identifiers = FALSE)

# Create network data
network_data <- list()

# Nodes: items and studies
item_nodes <- df %>%
  select(id = item_id, label = item_id, domain, title = stem_combined) %>%
  mutate(
    group = domain,
    type = "item",
    title = paste0("Item: ", id, "\nDomain: ", domain, "\nText: ", substr(title, 1, 100), "...")
  )

study_nodes <- df %>%
  separate_rows(studies, sep = ";") %>%
  filter(studies != "") %>%
  distinct(studies) %>%
  transmute(
    id = studies,
    label = studies,
    group = "study",
    type = "study",
    title = paste0("Study: ", studies)
  )

# Edges: item-study connections
edges <- df %>%
  separate_rows(studies, sep = ";") %>%
  filter(studies != "") %>%
  transmute(
    from = item_id,
    to = studies,
    title = paste0(item_id, " → ", studies)
  )

network_data$nodes <- bind_rows(item_nodes, study_nodes)
network_data$edges <- edges

# Treemap data
treemap_data <- df %>%
  separate_rows(studies, sep = ";") %>%
  filter(studies != "") %>%
  count(domain, studies, name = "value") %>%
  arrange(domain, studies)

# Sunburst data for plotly
sunburst_data <- df %>%
  separate_rows(studies, sep = ";") %>%
  filter(studies != "") %>%
  count(domain, studies, name = "value")

# Embed data as JSON
cat("<script>")
cat("window.networkData = ", toJSON(network_data, auto_unbox = TRUE), ";")
cat("window.treemapData = ", toJSON(treemap_data, auto_unbox = TRUE), ";")
cat("window.sunburstData = ", toJSON(sunburst_data, auto_unbox = TRUE), ";")
cat("window.codebookData = ", toJSON(codebook, auto_unbox = TRUE), ";")
cat("</script>")
```

<script src="https://unpkg.com/vis-network/standalone/umd/vis-network.min.js"></script>
<script src="https://cdn.plot.ly/plotly-latest.min.js"></script>

<script>
// Color scheme for domains
const domainColors = {
  'socemo': '#FF6B6B',
  'motor': '#4ECDC4',
  'coglan': '#45B7D1',
  'early_learning': '#96CEB4',
  'self_regulation': '#FECA57',
  'study': '#95A5A6'
};

// Initialize all visualizations
document.addEventListener('DOMContentLoaded', function() {
  initializeSunburst();
  initializeNetwork();
  initializeTreemap();
  populateFilters();
  setupEventHandlers();
});

function initializeSunburst() {
  const data = prepareSunburstData(1);

  const layout = {
    title: 'Codebook Hierarchy',
    font: { size: 12 },
    margin: { l: 0, r: 0, t: 50, b: 0 }
  };

  Plotly.newPlot('sunburst-chart', data, layout, {responsive: true});
}

function prepareSunburstData(level) {
  const data = window.sunburstData;
  let labels = ['Codebook'];
  let parents = [''];
  let values = [data.reduce((sum, d) => sum + d.value, 0)];
  let colors = ['#2E86AB'];

  if (level >= 1) {
    // Add domains
    const domainData = data.reduce((acc, d) => {
      if (!acc[d.domain]) acc[d.domain] = 0;
      acc[d.domain] += d.value;
      return acc;
    }, {});

    Object.entries(domainData).forEach(([domain, value]) => {
      labels.push(domain);
      parents.push('Codebook');
      values.push(value);
      colors.push(domainColors[domain] || '#95A5A6');
    });
  }

  if (level >= 2) {
    // Add studies within domains
    data.forEach(d => {
      labels.push(`${d.studies} (${d.domain})`);
      parents.push(d.domain);
      values.push(d.value);
      colors.push(domainColors[d.domain] || '#95A5A6');
    });
  }

  return [{
    type: 'sunburst',
    labels: labels,
    parents: parents,
    values: values,
    marker: { colors: colors },
    branchvalues: 'total',
    hovertemplate: '<b>%{label}</b><br>Items: %{value}<extra></extra>',
    textinfo: 'label+percent parent'
  }];
}

function initializeNetwork() {
  const container = document.getElementById('network-chart');

  // Prepare nodes with colors
  const nodes = window.networkData.nodes.map(node => ({
    ...node,
    color: {
      background: domainColors[node.group] || '#95A5A6',
      border: '#2B2B2B',
      highlight: {
        background: domainColors[node.group] || '#95A5A6',
        border: '#2B2B2B'
      }
    },
    shape: node.type === 'study' ? 'box' : 'dot',
    size: node.type === 'study' ? 25 : 15,
    font: { color: '#2B2B2B', size: 12 }
  }));

  const edges = window.networkData.edges.map(edge => ({
    ...edge,
    color: { color: '#848484', highlight: '#2B2B2B' },
    width: 1
  }));

  const data = { nodes: new vis.DataSet(nodes), edges: new vis.DataSet(edges) };

  const options = {
    physics: {
      enabled: true,
      forceAtlas2Based: {
        gravitationalConstant: -26,
        centralGravity: 0.005,
        springLength: 230,
        springConstant: 0.18
      },
      maxVelocity: 146,
      solver: 'forceAtlas2Based',
      timestep: 0.35,
      stabilization: { iterations: 150 }
    },
    nodes: {
      borderWidth: 2,
      shadow: true
    },
    edges: {
      shadow: true,
      smooth: {
        type: 'continuous'
      }
    },
    interaction: {
      hover: true,
      tooltipDelay: 200
    }
  };

  window.network = new vis.Network(container, data, options);

  // Add click handler
  window.network.on('click', function(params) {
    if (params.nodes.length > 0) {
      const nodeId = params.nodes[0];
      const node = nodes.find(n => n.id === nodeId);

      if (node && node.type === 'item') {
        showItemDetails(nodeId);
      }
    }
  });
}

function initializeTreemap() {
  const data = prepareTreemapData('domain-study');

  const layout = {
    title: 'Items by Domain and Study',
    font: { size: 12 },
    margin: { l: 0, r: 0, t: 50, b: 0 }
  };

  Plotly.newPlot('treemap-chart', data, layout, {responsive: true});
}

function prepareTreemapData(grouping) {
  const data = window.treemapData;

  if (grouping === 'domain-study') {
    return [{
      type: 'treemap',
      labels: data.map(d => `${d.studies}<br>(${d.value} items)`),
      parents: data.map(d => d.domain),
      values: data.map(d => d.value),
      text: data.map(d => `${d.studies}: ${d.value} items`),
      textinfo: 'label+text',
      hovertemplate: '<b>%{label}</b><br>Domain: %{parent}<br>Items: %{value}<extra></extra>',
      marker: {
        colors: data.map(d => domainColors[d.domain] || '#95A5A6')
      }
    }];
  }

  // Add other grouping options as needed
  return [];
}

function populateFilters() {
  // Populate network domain filter
  const domains = [...new Set(window.networkData.nodes
    .filter(n => n.type === 'item')
    .map(n => n.group))];

  const domainSelect = document.getElementById('network-filter-domain');
  domains.forEach(domain => {
    const option = document.createElement('option');
    option.value = domain;
    option.textContent = domain;
    domainSelect.appendChild(option);
  });
}

function setupEventHandlers() {
  // Sunburst controls
  document.getElementById('sunburst-level').addEventListener('change', function() {
    const level = parseInt(this.value);
    const data = prepareSunburstData(level);
    Plotly.react('sunburst-chart', data);
  });

  // Network controls
  document.getElementById('network-filter-domain').addEventListener('change', function() {
    filterNetworkByDomain(this.value);
  });

  document.getElementById('network-layout').addEventListener('change', function() {
    updateNetworkLayout(this.value);
  });

  document.getElementById('show-labels').addEventListener('change', function() {
    toggleNetworkLabels(this.checked);
  });

  // Treemap controls
  document.getElementById('treemap-grouping').addEventListener('change', function() {
    const data = prepareTreemapData(this.value);
    Plotly.react('treemap-chart', data);
  });
}

function filterNetworkByDomain(domain) {
  if (!window.network) return;

  const allNodes = window.networkData.nodes;
  const allEdges = window.networkData.edges;

  if (!domain) {
    // Show all
    window.network.setData({
      nodes: new vis.DataSet(allNodes),
      edges: new vis.DataSet(allEdges)
    });
    return;
  }

  // Filter nodes
  const filteredNodes = allNodes.filter(node =>
    node.type === 'study' || node.group === domain
  );

  // Filter edges
  const nodeIds = new Set(filteredNodes.map(n => n.id));
  const filteredEdges = allEdges.filter(edge =>
    nodeIds.has(edge.from) && nodeIds.has(edge.to)
  );

  window.network.setData({
    nodes: new vis.DataSet(filteredNodes),
    edges: new vis.DataSet(filteredEdges)
  });
}

function updateNetworkLayout(layout) {
  if (!window.network) return;

  const options = {
    physics: {
      solver: layout,
      stabilization: { iterations: 150 }
    }
  };

  window.network.setOptions(options);
}

function toggleNetworkLabels(show) {
  if (!window.network) return;

  const options = {
    nodes: {
      font: { color: show ? '#2B2B2B' : 'transparent' }
    }
  };

  window.network.setOptions(options);
}

function showItemDetails(itemId) {
  const item = window.codebookData.items[itemId];
  if (!item) return;

  const details = `
    <div class="modal fade" id="itemModal" tabindex="-1">
      <div class="modal-dialog modal-lg">
        <div class="modal-content">
          <div class="modal-header">
            <h5 class="modal-title">Item: ${itemId}</h5>
            <button type="button" class="btn-close" data-bs-dismiss="modal"></button>
          </div>
          <div class="modal-body">
            <p><strong>Domain:</strong> ${item.classification?.domain || 'Unknown'}</p>
            <p><strong>Studies:</strong> ${item.studies ? item.studies.join(', ') : 'None'}</p>
            <p><strong>Text:</strong> ${item.content?.stems?.combined || 'No text available'}</p>
            <hr>
            <pre class="language-json"><code>${JSON.stringify(item, null, 2)}</code></pre>
          </div>
        </div>
      </div>
    </div>
  `;

  // Remove existing modal if any
  const existingModal = document.getElementById('itemModal');
  if (existingModal) existingModal.remove();

  // Add new modal
  document.body.insertAdjacentHTML('beforeend', details);

  // Show modal
  const modal = new bootstrap.Modal(document.getElementById('itemModal'));
  modal.show();
}
</script>