---
title: "Variable-Project Matrix"
subtitle: "Which variables appear in which REDCap projects"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Ensure required packages are available
source("assets/ensure_dependencies.R")

# Load required libraries
library(dplyr)
library(DT)
library(knitr)

# Source the dictionary functions (now JSON-based)
source("assets/dictionary_functions.R")

# Initialize dictionary data from JSON
dict_data <- initialize_dictionary()

# Load data from JSON
matrix_data <- create_variable_project_matrix(dict_data)
```

# Variable-Project Matrix

```{r freshness-warning, results='asis', echo=FALSE}
# Display freshness warning if data is stale
cat(display_freshness_warning(dict_data))
```

This matrix shows which variables appear in which REDCap projects. Variables are identified by their field name, and checkmarks (✓) indicate presence in each project.

## Project Information

| PID | Description |
|-----|-------------|
| **7679** | Kidsights Data Survey |
| **7943** | Kidsights Email Registration |
| **7999** | Kidsights Public |
| **8014** | Kidsights Public Birth |

## Interactive Matrix

The table below shows all unique variables and their presence across the four REDCap projects. Use the search box to find specific variables, or sort by clicking column headers.

```{r variable-project-matrix}
# Display the variable-project matrix
if (nrow(matrix_data) > 0) {
  format_variable_project_matrix(matrix_data)
} else {
  cat("No matrix data available in JSON.")
}
```

## Summary Statistics

```{r matrix-summary}
# Get project coverage summary from JSON
summaries <- get_summary_data(dict_data)

if ("project_coverage" %in% names(summaries)) {
  project_summary <- summaries$project_coverage$data

  # Convert to data frame if needed
  if (is.list(project_summary)) {
    project_summary <- do.call(rbind, lapply(project_summary, data.frame))
  }

  knitr::kable(project_summary,
               col.names = c("Project ID", "Unique Variables", "Unique Forms"),
               caption = "Variable and Form Counts by Project")
}

# Calculate matrix statistics if data available
if (nrow(matrix_data) > 0 && ncol(matrix_data) >= 6) {
  total_vars <- nrow(matrix_data)

  # Check if we have the expected PID columns
  pid_cols <- grep("^PID_", names(matrix_data), value = TRUE)

  if (length(pid_cols) > 0) {
    # Calculate cross-project presence
    vars_in_all <- sum(rowSums(matrix_data[, pid_cols, drop = FALSE]) == length(pid_cols))
    vars_in_one <- sum(rowSums(matrix_data[, pid_cols, drop = FALSE]) == 1)
    vars_in_multiple <- sum(rowSums(matrix_data[, pid_cols, drop = FALSE]) > 1)

    # Create summary table
    summary_stats <- data.frame(
      Statistic = c("Total unique variables",
                    paste("Variables in all", length(pid_cols), "projects"),
                    "Variables in only 1 project",
                    "Variables in multiple projects"),
      Count = c(total_vars, vars_in_all, vars_in_one, vars_in_multiple),
      Percentage = round(c(100,
                          vars_in_all/total_vars*100,
                          vars_in_one/total_vars*100,
                          vars_in_multiple/total_vars*100), 1)
    )

    knitr::kable(summary_stats,
                 col.names = c("Statistic", "Count", "Percentage (%)"),
                 caption = "Variable Distribution Across Projects")
  }
}
```

```{r generation-info, echo=FALSE}
# Display generation information
if ("metadata" %in% names(dict_data)) {
  cat("\n**Data Dictionary Generated:** ", dict_data$metadata$generated, "\n")
  cat("**Total Raw Variables:** ", dict_data$metadata$total_raw_variables, "\n")
  cat("**Projects:** ", paste(dict_data$metadata$project_pids, collapse = ", "), "\n")
}
```